---
title: "Part 2 Chi Square with Lizards Tails"
author: "Carly Norris"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```


## Read in data

```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```


Get counts of lizards by species
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```
## Make a subset


only keep:
- UTST lizards
- three sites with the highest trap populations sites: IBPE (enclosure grassland), RABB(rabbit mesquite site), "EAST" (east tarbrush site)
- tail condition is recorded (W = whole), B= (broken)

```{r}
utst <- lizards %>% 
  filter(spp== "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```
- can use unique function to assure that what you have restricted in the function is actually occurring

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

nice to display both of these tables
```{r}
# table with counts
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

# table with proportions by row (default is by rows), all in janitor package
tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2)%>% # makes it a percent and rounds to two digits  
  adorn_ns() # adds the counts to the table 
```

Is there a sig association between site and tail breakage? Can we reject the null that tail breakage is independent of site?

## Convert to a contingency table
Need to only have counts in the table to run chi- square. can convert column to row names using 'column_to_rownames()'
 
```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

Are tail breakage and site independent?

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2) #used for inline referencing
```
Can reject the null: there is a significant association between site and tail breakage

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`)






